---
title: "Differential Gene Expression Analysis"
subtitle: "211001_APOE_Mm dataset"
author: "Lachlan Baer"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    fig_width: 8
    fig_height: 6
    fig_align: "center"
    code_folding: hide
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  error = FALSE,
  fig.align = "center"
)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

# Setup

```{r packages}
suppressPackageStartupMessages({
  ## Common
  library(tidyverse)
  library(magrittr)
  library(future.apply)
  library(here)
  library(AnnotationHub)
  library(purrr)
  library(scales)
  library(kableExtra)
  library(tictoc)
  library(ggrepel)
  library(RColorBrewer)
  library(ggpubr)
  library(pander)
  library(rmarkdown)
  ## Project specific
  library(edgeR)
  library(fgsea)
  library(pheatmap)
  library(cqn)
  library(DT)
  library(htmltools)
  library(gtools)
})
```

```{r options}
if (interactive()) setwd(here::here())
theme_set(theme_bw())
cores <- availableCores() - 1
```

```{r funcs}
source("~/bioinformatics/bioToolkit/lbFuncs.R")
```

## EnsDb

```{r ensParams}
ens_species <- "Mus musculus"
ens_release <- "104"
ens_assembly <- "GRCm39"
```

```{r ensDb}
ah <- AnnotationHub() %>%
  subset(species == ens_species) %>%
  subset(rdataclass == "EnsDb")
ahId <- ah$ah_id[str_detect(ah$title, ens_release)]
ensDb <- ah[[ahId]]
```

```{r chrInfo}
chrInfo <- getChromInfoFromEnsembl(ens_assembly, release = ens_release) %>%
  dplyr::filter(coord_system == "chromosome")
## Sort by numbers then letters
primary_chrs <- chrInfo$name %>%
  mixedsort()
## Push MT chr to end
primary_chrs <- c(
  primary_chrs[primary_chrs != "MT"],
  primary_chrs[primary_chrs == "MT"]
)
```

```{r transcripts}
transcripts <- transcripts(ensDb, filter = SeqNameFilter(primary_chrs))
txLen <- exonsBy(ensDb, "tx", filter = SeqNameFilter(primary_chrs)) %>%
  width() %>%
  vapply(sum, integer(1)) %>%
  enframe(name = "tx_id", value = "length")
mcols(transcripts) <- mcols(transcripts)[
  c("tx_id", "gene_id", "gc_content")
] %>%
  as.data.frame() %>%
  left_join(txLen)
```

```{r geneGc}
geneGc <- transcripts %>%
  mcols() %>%
  as_tibble() %>%
  group_by(gene_id) %>%
  summarise(
    gc_content = sum(gc_content*length) / sum(length),
    length = ceiling(median(length))
  )
```

```{r genes}
genes <- genes(ensDb, filter = SeqNameFilter(primary_chrs))
mcols(genes) <- mcols(genes)[
  c("gene_id", "gene_name", "gene_biotype", "entrezid")
] %>%
  as.data.frame() %>%
  left_join(geneGc)
```

An `EnsDb` object was obtained for Ensembl release `r ens_release` with the `AnnotationHub` package.
This contained the information required to set up gene and transcript annotations.
Gene-level estimates of GC content and length were also generated which are required for assessing potential biases in downstream analysis.

## Metadata

```{r metadata}
metadata <- read_tsv(here("misc/SYNAPSE_METADATA_MANIFEST.tsv")) %>%
  left_join(read_csv(here("misc/metaAPOE.csv"))) %>%
  dplyr::select(
    sample = specimenID, species, genotypeBackground, litter, dateBirth,
    dateDeath, genotype = Genotype, sex = Sex, age = Age, lane, basename = name,
    modelSystemName, individualID, study
  ) %>%
  dplyr::filter(str_detect(sample, "_3M_")) %>%
  mutate(basename = str_remove(basename, ".bam_R(1|2).fastq.gz")) %>%
  distinct(sample, .keep_all = TRUE) %>%
  mutate(
    group = as.factor(paste0(genotype, "_", age, "_", sex)),
    genotype = as.factor(genotype)
  ) %>%
  dplyr::arrange(genotype, group)
```

```{r genoCols}
genoCols <- metadata$genotype %>%
  unique() %>%
  length() %>%
  brewer.pal("Set1") %>%
  setNames(unique(metadata$genotype))
```

## Counts

```{r counts}
counts <- read_tsv(here("analysis-de/results/03_counts/counts.out"), comment = "#") %>%
  dplyr::select(-c(Chr, Start, End, Strand, Length)) %>%
  set_colnames(basename(colnames(.))) %>%
  set_colnames(str_remove(colnames(.), "\\.FCHJKY5BBXX.+"))
```

```{r dgeList}
minLib <- counts %>%
  dplyr::select(-Geneid) %>%
  colSums() %>%
  min()
# minCPM <- as.vector(cpm(10, lib.size = minLib))
minCPM <- 1
minSamples <- metadata %>%
  group_by(genotype) %>%
  summarise(n = n()) %>%
  pull(n) %>%
  min()
dgeList <- counts %>%
  as.data.frame() %>%
  column_to_rownames("Geneid") %>%
  .[rowSums(cpm(.) >= minCPM) >= minSamples,] %>%
  DGEList(
    samples = tibble(sample = colnames(.)) %>%
      left_join(metadata),
    genes = genes[rownames(.),] %>%
      as_tibble() %>%
      dplyr::select(chromosome = seqnames, everything()) %>%
      mutate(chromosome = factor(chromosome, levels = primary_chrs))
  ) %>%
  calcNormFactors()
```

```{r save_dgeList}
dgeList_path <- here("files/dgeList.Rds")
if (!file.exists(dgeList_path)) {
  saveRDS(dgeList, dgeList_path)
}
```

Gene-level count data was loaded from `featureCounts` and imported into `R` as a `DGEList` object.
Genes were removed from the counts matrix if they had less than `r round(minCPM, digits = 2)` CPM (~10 reads) in more than `r ncol(dgeList) - minSamples` samples.
This meant that each remaining gene must have assigned reads detected in at least one genotype group.

`r nrow(dgeList)` genes remained for further analysis with library sizes ranging from `r comma(min(dgeList$samples$lib.size))` to `r comma(max(dgeList$samples$lib.size))`.

```{r, fig.cap="*Expression density plots for all samples after filtering*"}
cpm(dgeList, log = TRUE) %>%
  as.data.frame() %>%
  pivot_longer(
    cols = everything(),
    names_to = "sample",
    values_to = "logCPM"
  ) %>%
  left_join(metadata) %>%
  ggplot(aes(logCPM, colour = genotype, group = sample)) +
  geom_density() +
  scale_colour_manual(values = genoCols) +
  labs(
    x = "logCPM",
    y = "Density",
    colour = "Genotype"
  )
```

# Initial DE analysis

## PCA

```{r}
cpm <- dgeList %>%
  cpm(log = TRUE)
```

```{r}
pca <- cpm %>%
  t() %>%
  prcomp()
pcaVars <- percent_format(0.1)(summary(pca)$importance["Proportion of Variance",])
```

```{r, fig.cap="*Principal Component Analysis of gene-level counts*"}
pca$x %>%
  as.data.frame() %>%
  rownames_to_column("sample") %>%
  left_join(metadata) %>%
  as_tibble() %>%
  ggplot(aes(PC1, PC2, colour = genotype, fill = genotype, label = sample)) +
  geom_point(size = 2) +
  geom_text_repel(show.legend = FALSE) +
  stat_ellipse(geom = "polygon", alpha = 0.05, show.legend = FALSE) +
  guides(fill = "none") +
  scale_colour_manual(values = genoCols) +
  labs(
    x = paste0("PC1 (", pcaVars[["PC1"]], ")"),
    y = paste0("PC2 (", pcaVars[["PC2"]], ")"),
    colour = "Genotype"
  )
```

## Design

```{r design}
design <- model.matrix(~0 + group, data = dgeList$samples) %>%
  set_colnames(str_remove(colnames(.), "group"))
```

```{r}
contrasts <- makeContrasts(
  APOE2v3_female = APOE2_3M_female - APOE3_3M_female,
  APOE2v3_male = APOE2_3M_male - APOE3_3M_male,
  APOE4v3_female = APOE4_3M_female - APOE3_3M_female,
  APOE4v3_male = APOE4_3M_male - APOE3_3M_male,
  levels = colnames(design)
)
```

```{r, fig.height=6, fig.width=6, fig.cap = "*Visualisation of the design matrix*"}
pheatmap(
  design[metadata$sample,],
  cluster_cols = FALSE,
  cluster_rows = FALSE,
  color = c("white", "grey50"),
  annotation_row = dgeList$samples["genotype"],
  annotation_colors = list(genotype = genoCols),
  angle_col = "315",
  legend = FALSE
)
```

## Model fitting

Genewise negative binomial generalized linear models were using the `glmFit()` function of the `edgeR` package.
Likelihood ratio tests were then performed for the coefficients in the model with `glmLRT()`.
Here the null hypothesis, $H_0$, is testing that the true values of the coefficients are equal to 0.
Due to the subtle effects of the mutation outlined in the PCA, genes were considered to be DE if they had an FDR-adjusted p-value < 0.05, as opposed to more stringent error rate controls such as the Bonferroni method.

```{r}
alpha <- 0.05
minLfc <- 1
fit <- estimateDisp(dgeList, design) %>%
  glmFit()
topTables <- colnames(contrasts) %>%
  sapply(function(x){
    glmLRT(fit, contrast = contrasts[,x]) %>%
      topTags(n = Inf) %>%
      .[["table"]] %>%
      as_tibble() %>%
      arrange(PValue) %>%
      dplyr::select(
        gene_id, gene_name, logFC, logCPM, PValue, FDR, everything()
      ) %>%
      mutate(
        contrast = x,
        bonfP = p.adjust(PValue, "bonf"),
        DE = bonfP < alpha
        # DE = case_when(
        #   bonfP < alpha ~ TRUE,
        #   FDR < alpha & abs(logFC) > minLfc ~ TRUE
        # ),
        # DE = ifelse(is.na(DE), FALSE, DE)
      )
  }, simplify = FALSE)
```

With this criteria, the following genes were classified as DE:

```{r}
for (ind in seq_along(topTables)) {
  print({
    topTables[[ind]] %>%
      dplyr::filter(DE) %>%
      dplyr::select(gene_id, gene_name, chromosome, logFC, logCPM, PValue, FDR, bonfP) %>%
      mutate(
        logCPM = formatC(logCPM, digits = 2, format = "f"),
        logFC = formatC(logFC, digits = 2, format = "f"),
        PValue = formatC(PValue, digits = 2, format = "e"),
        FDR = formatC(FDR, digits = 2, format = "e"),
        bonfP = formatC(bonfP, digits = 2, format = "e"),
      ) %>%
      kable(
        align = "l",
        caption = paste0(names(topTables[ind]), ". The ", nrow(.), " differentially expressed genes.")
      ) %>%
      kable_styling(
        bootstrap_options = c("striped", "hover", "condensed", "responsive")
      )
  })
}
```

```{r}
deCols <- c(`TRUE` = "red", `FALSE` = "grey50")
```

## Bias checks

### MA plots

```{r, fig.cap="*MA plots checking for logFC artefacts across the range of expression values. The average logFC appears consistant across all expression values for both comparisons*"}
topTables %>%
  bind_rows() %>%
  arrange(DE) %>%
  ggplot(aes(logCPM, logFC)) +
  geom_point(aes(colour = DE), alpha = 0.4) +
  # geom_text_repel(
  #   aes(label = gene_name, colour = DE),
  #   data = . %>% dplyr::filter(DE),
  #   max.overlaps = 20
  # ) +
  geom_smooth(se = FALSE) +
  facet_wrap(~contrast, ncol= 2) +
  scale_y_continuous(breaks = seq(-8, 8, by = 2)) +
  scale_colour_manual(values = deCols) +
  theme(legend.position = "none")
```

### GC content

```{r, fig.cap="*GC content bias for differential expression. GC content is shown against the ranking statistic, calculated by -log10(p) multiple by the sign of logFC. Differentially expressed genes are labelled red. Small biases were observed for both comparisons. In particular, the fAI genotype shows a well-documented pattern of GC bias, whereby both low and high GC proportions are affected.*"}
topTables %>%
  bind_rows() %>%
  mutate(stat = -sign(logFC)*log10(PValue)) %>%
  ggplot(aes(gc_content, stat)) +
  geom_point(aes(colour = DE), alpha = 0.4) +
  geom_smooth(se = FALSE) +
  facet_wrap(~contrast, ncol = 2)  +
  # geom_text_repel(
  #   aes(label = gene_name, colour = DE),
  #   data = . %>% dplyr::filter(DE)
  # ) +
  labs(
    x = " Gene GC content (%)",
    y = "Ranking Statistic"
  ) +
  coord_cartesian(ylim = c(-10, 10)) +
  scale_colour_manual(values = deCols) +
  theme(legend.position = "none")
```

### Gene length

```{r, fig.cap="*Gene length bias for differential expression. Gene length is shown against the ranking statistic, calculated by -log10(p) multiple by the sign of logFC. Differentially expressed genes are labelled red. Minimal bias is observed for the fAD genotype, but the fAI genotype shows a bias for larger genes.*"}
topTables %>%
  bind_rows() %>%
  mutate(stat = -sign(logFC)*log10(PValue)) %>%
  ggplot(aes(length, stat)) +
  geom_point(aes(colour = DE), alpha = 0.4) +
  geom_smooth(se = FALSE) +
  facet_wrap(~contrast, ncol = 2)  +
  # geom_text_repel(
  #   aes(label = gene_name, colour = DE),
  #   data = . %>% dplyr::filter(DE)
  # ) +
  labs(
    x = "GC length",
    y = "Ranking Statistic"
  ) +
  coord_cartesian(ylim = c(-10, 10)) +
  scale_x_log10(labels = comma) +
  scale_colour_manual(values = deCols) +
  theme(legend.position = "none")
```

# Conditional Quantile Normalisation

Conditional Quantile Normalisation (CQN) is a procedure that can account for GC content and length biases in differential expression and is implemented with package `cqn`.
It does this by making by calculating offset values that can be incorporated into the GLM.
Both GC content and length were noted as being of concern and were therefore used to calculate offsets.
The resulting `glm.offset` values from the output of `cqn` were added to the original `DGEList` object.

```{r, results='hide'}
# cqn <- cqn(
#   counts = dgeList$counts,
#   x = dgeList$genes$gc_content,
#   lengths = rep(1000, times = nrow(dgeList)),
#   sizeFactors = dgeList$samples$lib.size,
#   lengthMethod = "fixed"
# )
cqn <- cqn(
  counts = dgeList$counts,
  x = dgeList$genes$gc_content,
  lengths = dgeList$genes$length,
  sizeFactors = dgeList$samples$lib.size
)
```

```{r, fig.cap="*Model fits for GC content and gene length under the CQN model. *"}
par(mfrow = c(1, 2))
cqnplot(cqn, n = 1, xlab = "GC Content", col = genoCols)
cqnplot(cqn, n = 2, xlab = "Length", col = genoCols)
legend(
  "bottomright",
  legend = levels(metadata$genotype),
  col = genoCols,
  lty = 1
)
par(mfrow = c(1, 1))
```

```{r}
dgeList$offset <- cqn$glm.offset
```

## PCA

```{r}
cpm_cqn <- cqn %>%
  with(y + offset)
```

```{r}
pca_cqn <- cpm_cqn %>%
  t() %>%
  prcomp() 
pcaVars <- percent_format(0.1)(summary(pca_cqn)$importance["Proportion of Variance",])
```

```{r, fig.cap="*Principal Component Analysis of gene-level counts post CQN.*"}
pca_cqn$x %>%
  as.data.frame() %>%
  rownames_to_column("sample") %>%
  left_join(metadata) %>%
  as_tibble() %>%
  ggplot(aes(PC1, PC2, colour = genotype, fill = genotype, label = sample)) +
  geom_point(size = 2) +
  geom_text_repel(show.legend = FALSE) +
  stat_ellipse(geom = "polygon", alpha = 0.05, show.legend = FALSE) +
  guides(fill = "none") +
  scale_colour_manual(values = genoCols) +
  labs(
    x = paste0("PC1 (", pcaVars[["PC1"]], ")"),
    y = paste0("PC2 (", pcaVars[["PC2"]], ")"),
    colour = "Genotype"
  )
```

There was very little improvement in separation of genotype groups after CQN, and PC1 still only accounted for `r pcaVars[["PC1"]]` of the variation.
The x-axis of the PCA plot appears to be flipped, but this normal as the signs of PC values are somewhat arbitrary.

## Model fitting

After incorporation of the offset values into the `DGEList` object, dispersion estimates were calculated and models were fitted as done pre CQN.

```{r}
alpha <- 0.05
minLfc <- 1
fit_cqn <- estimateDisp(dgeList, design) %>%
  glmFit()
topTables_cqn <- colnames(contrasts) %>%
  sapply(function(x){
    glmLRT(fit_cqn, contrast = contrasts[,x]) %>%
      topTags(n = Inf) %>%
      .[["table"]] %>%
      as_tibble() %>%
      arrange(PValue) %>%
      dplyr::select(
        gene_id, gene_name, logFC, logCPM, PValue, FDR, everything()  
      ) %>%
      mutate(
        contrast = x,
        bonfP = p.adjust(PValue, "bonf"),
        DE = bonfP < alpha
        # DE = case_when(
        #   bonfP < alpha ~ TRUE,
        #   FDR < alpha & abs(logFC) > minLfc ~ TRUE
        # ),
        # DE = ifelse(is.na(DE), FALSE, DE)
      )
  }, simplify = FALSE)
```


```{r}
lapply(seq_along(topTables_cqn), function(ind){
  topTables_cqn[[ind]] %>%
    dplyr::filter(DE) %>% 
    dplyr::select(gene_id, gene_name, chromosome, logFC, logCPM, PValue, FDR, bonfP) %>%
    mutate(
      logCPM = formatC(logCPM, digits = 2, format = "f"),
      logFC = formatC(logFC, digits = 2, format = "f"),
      PValue = formatC(PValue, digits = 2, format = "e"),
      FDR = formatC(FDR, digits = 2, format = "e"),
      bonfP = formatC(bonfP, digits = 2, format = "e"),
    ) %>%
    kable(
      align = "l",
      caption = paste0(names(topTables[ind]), ". The ", nrow(.), " differentially expressed genes.")
    ) %>%
    kable_styling(
      bootstrap_options = c("striped", "hover", "condensed", "responsive")
    )
})
```

## Bias checks

### MA plots

```{r, fig.cap="*MA plots checking for logFC artefacts across the range of expression values. The average logFC remains consistant across all expression values for both comparisons after CQN*"}
topTables_cqn %>%
  bind_rows() %>%
  arrange(DE) %>%
  ggplot(aes(logCPM, logFC)) +
  geom_point(aes(colour = DE), alpha = 0.4) +
  # geom_text_repel(
  #   aes(label = gene_name, colour = DE),
  #   data = . %>% dplyr::filter(DE)
  # ) +
  geom_smooth(se = FALSE) +
  facet_wrap(~contrast, ncol = 2) +
  scale_y_continuous(breaks = seq(-8, 8, by = 2)) +
  scale_colour_manual(values = deCols) +
  theme(legend.position = "none")
```

### GC content

```{r, fig.cap="*GC content bias for differential expression. GC content is shown against the ranking statistic, calculated by -log10(p) multiple by the sign of logFC. Differentially expressed genes are labelled red. No improvement of GC bias was observed post-CQN. In fact, GC bias appeared worse than pre-CQN.*"}
topTables_cqn %>%
  bind_rows() %>%
  mutate(stat = -sign(logFC)*log10(PValue)) %>%
  ggplot(aes(gc_content, stat)) +
  geom_point(aes(colour = DE), alpha = 0.4) +
  geom_smooth(se = FALSE) +
  facet_wrap(~contrast, ncol = 2)  +
  # geom_text_repel(
  #   aes(label = gene_name, colour = DE),
  #   data = . %>% dplyr::filter(DE)
  # ) +
  labs(
    x = "Gene GC content (%)",
    y = "Ranking Statistic"
  ) +
  coord_cartesian(ylim = c(-10, 10)) +
  scale_colour_manual(values = deCols) +
  theme(legend.position = "none")
```

### Gene length

```{r, fig.cap="*Gene length bias for differential expression. Gene length is shown against the ranking statistic, calculated by -log10(p) multiple by the sign of logFC. Differentially expressed genes are labelled red. No improvement of gene length bias was observed for either genotype post-CQN.*"}
topTables_cqn %>%
  bind_rows() %>%
  mutate(stat = -sign(logFC)*log10(PValue)) %>%
  ggplot(aes(length, stat)) +
  geom_point(aes(colour = DE), alpha = 0.4) +
  geom_smooth(se = FALSE) +
  facet_wrap(~contrast, ncol = 2)  +
  # geom_text_repel(
  #   aes(label = gene_name, colour = DE),
  #   data = . %>% dplyr::filter(DE)
  # ) +
  labs(
    x = "GC length",
    y = "Ranking Statistic"
  ) +
  coord_cartesian(ylim = c(-10, 10)) +
  scale_x_log10(labels = comma) +
  scale_colour_manual(values = deCols) +
  theme(legend.position = "none")
```

# Results

```{r}
topTables_chosen <- topTables_cqn
```

```{r, fig.cap="*Volcano plots showing -log10(p) against logFC. Genes determined to be DE are coloured red.*"}
topTables_chosen %>%
  bind_rows() %>%
  ggplot(aes(logFC, -log10(PValue), colour = DE)) +
  geom_point(alpha = 0.4) +
  # geom_text_repel(
  #   aes(label = gene_name),
  #   size = 3,
  #   data = . %>% dplyr::filter(DE)
  # ) +
  # geom_text_repel(
  #   aes(label = gene_name),
  #   size = 3,
  #   data = . %>% dplyr::filter(!DE, logFC < -3)
  # ) +
  facet_wrap(~contrast, ncol = 2) +
  scale_colour_manual(values = deCols) +
  scale_x_continuous(breaks = seq(-8, 8, by = 2)) +
  coord_cartesian(ylim = c(0, 10)) +
  labs(y = log10plab) +
  theme(legend.position = "none")
```

```{r}
tt <- topTables_chosen$APOE2v3_female
man <- tt %>%
  mutate(mid = (start + end) / 2) %>%
  dplyr::select(gene_id, gene_name, chromosome, mid, FDR) %>%
  dplyr::filter(!(chromosome %in% c("X", "Y", "MT"))) %>%
  group_by(chromosome) %>%
  summarise(chrLen = max(mid)) %>%
  mutate(chrSt = cumsum(chrLen)-chrLen) %>%
  dplyr::select(-chrLen) %>% 
  left_join(tt) %>%
  mutate(mid = (start + end) / 2) %>%
  mutate(midCum = chrSt + mid) %>%
  mutate(
    colour = ifelse(chromosome %in% primary_chrs[c(TRUE, FALSE)], "a", "b"),
    colour = ifelse(chromosome == "7", "c", colour)
  )
axis <- man %>%
  group_by(chromosome) %>%
  summarise(center = (max(midCum) + min(midCum)) / 2)
ggplot(man, aes(x = midCum, y = -log10(PValue))) +
  geom_point(
    data = man %>%
      dplyr::filter(!DE),
    aes(colour = colour),
    shape = 20,
    alpha = 1,
    size = 2
  ) +
  scale_colour_manual(values = c("grey50", "grey80", "red")) +
  geom_point(
    data = man %>%
      dplyr::filter(DE),
    aes(fill = colour),
    colour = "black",
    shape = 21,
    alpha = 1,
    size = 1.5
  ) +
  scale_fill_manual(values = c("grey50", "grey80", "red")) +
  # scale_shape_manual(values = c(20, 21)) +
  scale_x_continuous(label = axis$chromosome, breaks = axis$center) +
  scale_y_continuous(expand = c(0, 0)) +
  coord_cartesian(clip = "off") +
  labs(
    # title = paste0(unique(topTable$contrast)),
    x = "Chromosome",
    y = expression(paste(-log[10], "(p)"))
  ) +
  # geom_text_repel(
  #   data = dplyr::filter(man, DE),
  #   aes(label = gene_name, colour = colour),
  #   size = 2.5,
  #   max.overlaps = 50
  # ) +
  theme(
    legend.position="none",
    panel.border = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  )
```

## Data export

Final results were exported as Rds files for further analysis.

```{r}
saveRDS(topTables, here("files/topTables.Rds"))
saveRDS(topTables_cqn, here("files/topTables_cqn.Rds"))
```

```{r}
sessionInfo() %>%
  pander()
```
