---
title: "Enrichment Analysis"
subtitle: "211001_APOE_Mm dataset"
author: "Lachlan Baer"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    fig_width: 8
    fig_height: 6
    fig_align: "center"
    code_folding: hide
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  error = FALSE,
  fig.align = "center"
)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

# Setup

```{r packages}
suppressPackageStartupMessages({
  ## Common
  library(tidyverse)
  library(magrittr)
  library(future.apply)
  library(here)
  library(AnnotationHub)
  library(purrr)
  library(scales)
  library(kableExtra)
  library(tictoc)
  library(ggrepel)
  library(RColorBrewer)
  library(ggpubr)
  library(pander)
  library(rmarkdown)
  ## Project specific
  library(edgeR)
  library(limma)
  library(pheatmap)
  library(cqn)
  library(DT)
  library(htmltools)
  library(msigdbr)
  library(reactable)
})
```

```{r options}
if (interactive()) setwd(here::here())
theme_set(theme_bw())
cores <- availableCores() - 1
```

```{r}
source("~/bioinformatics/bioToolkit/lbFuncs.R")
```

## Annotations

```{r ensParams}
ens_species <- "Mus musculus"
ens_release <- "104"
ens_assembly <- "GRCm39"
```

```{r ensDb}
ah <- AnnotationHub() %>%
  subset(species == ens_species) %>%
  subset(rdataclass == "EnsDb")
ahId <- ah$ah_id[str_detect(ah$title, ens_release)]
ensDb <- ah[[ahId]]
```

```{r chrInfo}
chrInfo <- getChromInfoFromEnsembl(ens_assembly, release = ens_release) %>%
  dplyr::filter(coord_system == "chromosome")
primary_chrs <- chrInfo$name
```

```{r transcripts}
transcripts <- transcripts(ensDb, filter = SeqNameFilter(primary_chrs))
txLen <- exonsBy(ensDb, "tx", filter = SeqNameFilter(primary_chrs)) %>%
  width() %>%
  vapply(sum, integer(1))
mcols(transcripts) <- mcols(transcripts)[
  c("tx_id", "gene_id", "gc_content")
] %>%
  as.data.frame() %>%
  mutate(length = txLen)
```

```{r geneGc}
geneGc <- transcripts %>%
  mcols() %>%
  as_tibble() %>%
  group_by(gene_id) %>%
  summarise(
    gc_content = sum(gc_content*length) / sum(length),
    length = ceiling(median(length))
  )
```

```{r genes}
genes <- genes(ensDb, filter = SeqNameFilter(primary_chrs)) %>%
  .[,c("gene_id", "gene_name", "gene_biotype", "entrezid")] %>%
  as_tibble() %>%
  left_join(geneGc) %>%
  GRanges()
```

An `EnsDb` object was obtained for Ensembl release 101 with the `AnnotationHub` package.
This contained the information required to set up gene and transcript annotations.
Gene-level estimates of GC content and length were also generated which may be required as covariates in suitable analyses.

```{r entrezGenes}
entrezGenes <- genes %>%
  as.data.frame() %>%
  dplyr::filter(!is.na(entrezid)) %>%
  mutate(entrezid = unAsIs(entrezid)) %>%
  unnest(entrezid) %>%
  dplyr::rename(entrez_gene = entrezid)
```

```{r geneChrs}
geneChrs <- as_tibble(genes) %>%
  dplyr::select(gene_id, chromosome = seqnames)
```

## Metadata

```{r metadata}
metadata <- read_tsv(here("misc/SYNAPSE_METADATA_MANIFEST.tsv")) %>%
  left_join(read_csv(here("misc/metaAPOE.csv"))) %>%
  dplyr::select(
    sample = specimenID, species, genotypeBackground, litter, dateBirth,
    dateDeath, genotype = Genotype, sex = Sex, age = Age, lane, basename = name,
    modelSystemName, individualID, study
  ) %>%
  dplyr::filter(str_detect(sample, "_3M_")) %>%
  mutate(basename = str_remove(basename, ".bam_R(1|2).fastq.gz")) %>%
  distinct(sample, .keep_all = TRUE) %>%
  mutate(
    group = as.factor(paste0(genotype, "_", age, "_", sex)),
    genotype = as.factor(genotype)
  ) %>%
  dplyr::arrange(genotype, group)
```

```{r genoCols}
genoCols <- metadata$genotype %>%
  unique() %>%
  length() %>%
  brewer.pal("Set1") %>%
  setNames(unique(metadata$genotype))
```

```{r samples_byGroup}
samples_byGroup <- metadata %>%
  split(f = .$group) %>%
  sapply(function(x){
    pull(x, sample)
  }, simplify = FALSE)
```

## Differential expression data

```{r dgeList}
dgeList <- readRDS(here("files/dgeList.Rds"))
```

```{r design}
design <- model.matrix(~0 + group, data = dgeList$samples) %>%
  set_colnames(str_remove(colnames(.), "group"))
```

```{r contrasts}
contrasts <- makeContrasts(
  APOE2v3_female = APOE2_3M_female - APOE3_3M_female,
  APOE2v3_male = APOE2_3M_male - APOE3_3M_male,
  APOE4v3_female = APOE4_3M_female - APOE3_3M_female,
  APOE4v3_male = APOE4_3M_male - APOE3_3M_male,
  levels = colnames(design)
)
```

```{r topTables}
topTables <- readRDS(here("files/topTables_cqn.Rds"))
```

```{r deGenes}
deGenes <- lapply(topTables, dplyr::filter, DE)
```

## Variant data

```{r diss_winRanges}
diss_winRanges <- readRDS(here("files/diss_winRanges.Rds")) %>%
  .[1:4]  # Only concerned with DE-tested comparisons
```

# Gene sets

Multiple gene set databases were accessed with the `msigdbr` package.
For this analysis the Hallmark, KEGG, Wikipathways and chromosomal based gene sets were chosen.

```{r hm}
hm <- msigdbr(ens_species, category = "H") %>%
  left_join(entrezGenes) %>%
  dplyr::filter(!is.na(gene_id)) %>%
  distinct(gs_name, gene_id, .keep_all = TRUE) %>%
  mutate(gs_name = str_remove(gs_name, "^HALLMARK_"))
hmByGene <- hm %>%
  split(f = .$gene_id) %>%
  lapply(extract2, "gs_name")
hmByID <- hm %>%
  split(f = .$gs_name) %>%
  lapply(extract2, "gene_id")
```

- The Hallmark gene sets consists of `r length(hmByID)` gene sets ranging from sizes of `r min(sapply(hmByID, length))` to `r max(sapply(hmByID, length))` genes.

```{r kg}
kg <- msigdbr(ens_species, category = "C2", subcategory = "CP:KEGG") %>%
  left_join(entrezGenes) %>%
  dplyr::filter(!is.na(gene_id)) %>%
  distinct(gs_name, gene_id, .keep_all = TRUE) %>%
  mutate(gs_name = str_remove(gs_name, "^KEGG_"))
kgByGene <- kg %>%
  split(f = .$gene_id) %>%
  lapply(extract2, "gs_name")
kgByID <- kg %>%
  split(f = .$gs_name) %>%
  lapply(extract2, "gene_id")
```

- The KEGG gene sets consists of `r length(kgByID)` gene sets ranging from sizes of `r min(sapply(kgByID, length))` to `r max(sapply(kgByID, length))` genes.

```{r wk}
wk <- msigdbr(ens_species, category = "C2", subcategory = "CP:WIKIPATHWAYS") %>%
  left_join(entrezGenes) %>%
  dplyr::filter(!is.na(gene_id)) %>%
  distinct(gs_name, gene_id, .keep_all = TRUE) %>%
  mutate(gs_name = str_remove(gs_name, "^WP_"))
wkByGene <- wk %>%
  split(f = .$gene_id) %>%
  lapply(extract2, "gs_name")
wkByID <- wk %>%
  split(f = .$gs_name) %>%
  lapply(extract2, "gene_id")
```

- The Wikipathways gene sets consists of `r length(wkByID)` gene sets ranging from sizes of `r min(sapply(wkByID, length))` to `r max(sapply(wkByID, length))` genes.

# Enrichment testing {.tabset}

## Hallmark {.tabset .tabset-pills}

```{r hm_fry}
hm_fry <- contrasts %>%
  apply(MARGIN = 2, function(contrast){
    cpm(dgeList$counts, log = TRUE) %>%
      fry(
        index = hmByID,
        design = design,
        contrast = contrast,
        sort = "mixed"
      ) %>%
      rownames_to_column("Pathway") %>%
      as_tibble() %>%
      mutate(
        Pathway = str_trunc(Pathway, width = 20),
        PValue = formatC(PValue, digits = 1, format = "e"),
        FDR = formatC(FDR, digits = 1, format = "e"),
        PValue.Mixed = formatC(PValue.Mixed, digits = 1, format = "e"),
        FDR.Mixed = formatC(FDR.Mixed, digits = 1, format = "e"),
      )
  })
```

### APOE 2v3 females

```{r}
datatable(hm_fry[[1]])
```

### APOE 2v3 males

```{r}
datatable(hm_fry[[2]])
```

### APOE 4v3 females

```{r}
datatable(hm_fry[[3]])
```

### APOE 4v3 males

```{r}
datatable(hm_fry[[4]])
```

## KEGG {.tabset .tabset-pills}

```{r kg_fry}
kg_fry <- contrasts %>%
  apply(MARGIN = 2, function(contrast){
    cpm(dgeList$counts, log = TRUE) %>%
      fry(
        index = kgByID,
        design = design,
        contrast = contrast,
        sort = "mixed"
      ) %>%
      rownames_to_column("Pathway") %>%
      as_tibble() %>%
      mutate(
        Pathway = str_trunc(Pathway, width = 20),
        PValue = formatC(PValue, digits = 1, format = "e"),
        FDR = formatC(FDR, digits = 1, format = "e"),
        PValue.Mixed = formatC(PValue.Mixed, digits = 1, format = "e"),
        FDR.Mixed = formatC(FDR.Mixed, digits = 1, format = "e"),
      )
  })
```

### APOE 2v3 females

```{r}
datatable(kg_fry[[1]])
```

### APOE 2v3 males

```{r}
datatable(kg_fry[[2]])
```

### APOE 4v3 females

```{r}
datatable(kg_fry[[3]])
```

### APOE 4v3 males

```{r}
datatable(kg_fry[[4]])
```

## Wikipathways {.tabset .tabset-pills}

```{r wk_fry}
wk_fry <- contrasts %>%
  apply(MARGIN = 2, function(contrast){
    cpm(dgeList$counts, log = TRUE) %>%
      fry(
        index = wkByID,
        design = design,
        contrast = contrast,
        sort = "mixed"
      ) %>%
      rownames_to_column("Pathway") %>%
      as_tibble() %>%
      mutate(
        Pathway = str_trunc(Pathway, width = 20),
        PValue = formatC(PValue, digits = 1, format = "e"),
        FDR = formatC(FDR, digits = 1, format = "e"),
        PValue.Mixed = formatC(PValue.Mixed, digits = 1, format = "e"),
        FDR.Mixed = formatC(FDR.Mixed, digits = 1, format = "e"),
      )
  })
```

### APOE 2v3 females

```{r}
datatable(wk_fry[[1]])
```

### APOE 2v3 males

```{r}
datatable(wk_fry[[2]])
```

### APOE 4v3 females

```{r}
datatable(wk_fry[[3]])
```

### APOE 4v3 males

```{r}
datatable(wk_fry[[4]])
```

# eQTL bias removal {.tabset}

```{r genes_diss_path}
genes_diss_path <- here("files/genes_diss.Rds")
```

```{r genes_diss}
## Takes 5 mins to run so .Rds saved for convenience
if (!file.exists(genes_diss_path)) {
  genes_diss <- lapply(diss_winRanges, function(winRanges){
    overlaps <- findOverlaps(GRanges(dgeList$genes), winRanges)
    lapply(unique(queryHits(overlaps)), function(query){
      subjects <- subjectHits(overlaps)[queryHits(overlaps) == query]
      diss <- winRanges$diss_rollmean[subjects] %>%
        mean()
      dgeList$genes[query,] %>%
        mutate(diss = diss)
    }) %>%
      bind_rows() %>%
      GRanges()
  })
  saveRDS(genes_diss, genes_diss_path)
} else {
  genes_diss <- readRDS(genes_diss_path)
}
```

## Hallmark {.tabset .tabset-pills}

```{r hm_fry_eqtl}
hm_fry_eqtl <- colnames(contrasts) %>%
  sapply(function(cont){
    biased_genes <- genes_diss[[cont]] %>%
      as_tibble() %>%
      dplyr::filter(diss > 0.5) %>%
      pull(gene_id)
    cpm(
      dgeList$counts[!(rownames(dgeList$counts) %in% biased_genes),],
      log = TRUE
    ) %>%
      fry(
        index = hmByID,
        design = design,
        contrast = contrasts[,cont],
        sort = "mixed"
      ) %>%
      rownames_to_column("Pathway") %>%
      as_tibble() %>%
      mutate(
        Pathway = str_trunc(Pathway, width = 20),
        PValue = formatC(PValue, digits = 1, format = "e"),
        FDR = formatC(FDR, digits = 1, format = "e"),
        PValue.Mixed = formatC(PValue.Mixed, digits = 1, format = "e"),
        FDR.Mixed = formatC(FDR.Mixed, digits = 1, format = "e"),
      )
  }, simplify = FALSE)
```

### APOE 2v3 females

```{r}
datatable(hm_fry_eqtl[[1]])
```

### APOE 2v3 males

```{r}
datatable(hm_fry_eqtl[[2]])
```

### APOE 4v3 females

```{r}
datatable(hm_fry_eqtl[[3]])
```

### APOE 4v3 males

```{r}
datatable(hm_fry_eqtl[[4]])
```

## KEGG {.tabset .tabset-pills}

```{r kg_fry_eqtl}
kg_fry_eqtl <- colnames(contrasts) %>%
  sapply(function(cont){
    biased_genes <- genes_diss[[cont]] %>%
      as_tibble() %>%
      dplyr::filter(diss > 0.5) %>%
      pull(gene_id)
    cpm(
      dgeList$counts[!(rownames(dgeList$counts) %in% biased_genes),],
      log = TRUE
    ) %>%
      fry(
        index = kgByID,
        design = design,
        contrast = contrasts[,cont],
        sort = "mixed"
      ) %>%
      rownames_to_column("Pathway") %>%
      as_tibble() %>%
      mutate(
        Pathway = str_trunc(Pathway, width = 20),
        PValue = formatC(PValue, digits = 1, format = "e"),
        FDR = formatC(FDR, digits = 1, format = "e"),
        PValue.Mixed = formatC(PValue.Mixed, digits = 1, format = "e"),
        FDR.Mixed = formatC(FDR.Mixed, digits = 1, format = "e"),
      )
  }, simplify = FALSE)
```

### APOE 2v3 females

```{r}
datatable(kg_fry_eqtl[[1]])
```

### APOE 2v3 males

```{r}
datatable(kg_fry_eqtl[[2]])
```

### APOE 4v3 females

```{r}
datatable(kg_fry_eqtl[[3]])
```

### APOE 4v3 males

```{r}
datatable(kg_fry_eqtl[[4]])
```

## Wikipathways {.tabset .tabset-pills}

```{r wk_fry_eqtl}
wk_fry_eqtl <- colnames(contrasts) %>%
  sapply(function(cont){
    biased_genes <- genes_diss[[cont]] %>%
      as_tibble() %>%
      dplyr::filter(diss > 0.5) %>%
      pull(gene_id)
    cpm(
      dgeList$counts[!(rownames(dgeList$counts) %in% biased_genes),],
      log = TRUE
    ) %>%
      fry(
        index = wkByID,
        design = design,
        contrast = contrasts[,cont],
        sort = "mixed"
      ) %>%
      rownames_to_column("Pathway") %>%
      as_tibble() %>%
      mutate(
        Pathway = str_trunc(Pathway, width = 20),
        PValue = formatC(PValue, digits = 1, format = "e"),
        FDR = formatC(FDR, digits = 1, format = "e"),
        PValue.Mixed = formatC(PValue.Mixed, digits = 1, format = "e"),
        FDR.Mixed = formatC(FDR.Mixed, digits = 1, format = "e"),
      )
  }, simplify = FALSE)
```

### APOE 2v3 females

```{r}
datatable(wk_fry_eqtl[[1]])
```

### APOE 2v3 males

```{r}
datatable(wk_fry_eqtl[[2]])
```

### APOE 4v3 females

```{r}
datatable(wk_fry_eqtl[[3]])
```

### APOE 4v3 males

```{r}
datatable(wk_fry_eqtl[[4]])
```
